Constraints on Cosmic-Ray Particle Populations, Magnetic Fields, and Dark Matter from VHE Gamma-Ray Observations of the Coma Cluster of Galaxies with VERITAS

### 2011-07-08 ms.tex
more conflicts resolved
Changes

\begin{abstract}
Theoretical models of galaxy clusters suggest they may be sources of very high-energy (VHE) gamma-ray emission. Observations of radio halos and relics in clusters indicate populations of relativistic electrons and positrons, which are either injected into the intra-cluster medium and accelerated by turbulent shocks or secondaries from interactions of cosmic-ray hadrons (protons and ions) with the intra-cluster gas. Those interactions would also produce gamma rays through the decay of neutral pions. Reported here are observations of the Coma cluster of galaxies with the VERITAS array of imaging Cherenkov telescopes. The observations were made between March and May 2008. Radio observations at 1.41 GHz taken with the Green Bank Telescope and X-ray observations with the ROSAT All-Sky survey was used to a priori select regions of interest in the field of view. No significant VHE gamma-ray emission from the Coma cluster was detected by VERITAS. Several emission regions are considered, including the cluster core as both unresolved (point-like) and spatially extended, and a select few cluster members. The flux upper limits at the 99\% confidence level measured are on the order of $2-5\times 10^{-8}$ ph. m$^{-2}$ s$^{-1}$, which is X times higher than the model-predicted minimum gamma-ray flux.

@@

%\keywords{cosmic rays: interactions --- galaxies --- galaxy clusters -- VHE}
\keywords{cosmic rays --- gamma rays: observations --- gamma rays: VHE --- galaxies: clusters: general --- galaxies: clusters: individual: Coma (ACO 1656)}

@@

\begin{deluxetable}{lcc}
\tablecolumns{3}
\tablewidth{0pc}
\tablecaption{Regions of interest in the Coma cluster field of view.}
\tablehead{
\colhead{Source} &
\colhead{RA} &
\colhead{Dec}}
\startdata
Core & \RA{12}{59}{48.7} & \Dec{+27}{58}{50.0}\\
NGC 4889 & \RA{13}{00}{08.13} & \Dec{+27}{58}{37.03}\\
NGC 4874 & \RA{12}{59}{35.71} & \Dec{+27}{57}{33.37}\\
NGC 4921 & \RA{13}{01}{26.12} & \Dec{+27}{53}{09.59}\\
\enddata
\tablenotetext{a}{Note. The cluster core is considered both as a point source and a modestly extended source. Three central galaxies are also considered in point-source searches. The choice is based on evidence of an excess of non-thermal X-ray emission \citep{article:Neumann_etal:2003} at the location of these galaxies.}
\label{table:roi}
\end{deluxetable}

@@

>>>>

\begin{abstract}
Theoretical models of galaxy clusters suggest they may be sources of very high-energy (VHE) gamma-ray emission. Observations of radio halos and relics in clusters indicate populations of relativistic electrons and positrons, which are either injected into the intra-cluster medium and accelerated by turbulent shocks or secondaries from interactions of cosmic-ray hadrons (protons and ions) with the intra-cluster gas. Those interactions would also produce gamma rays through the decay of neutral pions. Reported here are observations of the Coma cluster of galaxies with the VERITAS array of imaging Cherenkov telescopes. The observations were made between March and May 2008. Radio observations at 1.41 GHz taken with the Green Bank Telescope and X-ray observations with the ROSAT All-Sky survey was used to a priori select regions of interest in the field of view. No significant VHE gamma-ray emission from the Coma cluster was detected by VERITAS. Several emission regions are considered, including the cluster core as both unresolved (point-like) and spatially extended, and a select few cluster members. The measured flux upper limits at the 99\% confidence level are on the order of $2-5\times 10^{-8}$ ph. m$^{-2}$ s$^{-1}$, which is X times higher than the model-predicted minimum gamma-ray flux.

@@

%\keywords{cosmic rays: interactions --- galaxies --- galaxy clusters -- VHE}
\keywords{cosmic rays --- gamma rays: observations --- gamma rays: VHE --- galaxies: clusters: general --- galaxies: clusters: individual: Coma (ACO 1656) --- dark matter}

@@

\begin{deluxetable}{lcc}
\tablecolumns{3}
\tablewidth{0pc}
\tablecaption{Regions of interest in the Coma cluster field of view.}
\tablehead{
\colhead{Source} &
\colhead{RA} &
\colhead{Dec}}
\startdata
Core & \RA{12}{59}{48.7} & \Dec{+27}{58}{50.0}\\
NGC 4889 & \RA{13}{00}{08.13} & \Dec{+27}{58}{37.03}\\
NGC 4874 & \RA{12}{59}{35.71} & \Dec{+27}{57}{33.37}\\
NGC 4921 & \RA{13}{01}{26.12} & \Dec{+27}{53}{09.59}\\
\enddata
\tablecomments{The cluster core is considered both as a point source and a modestly extended source. Three central galaxies are also considered in point-source searches. The choice is based on evidence of an excess of non-thermal X-ray emission \citep{article:Neumann_etal:2003} at the location of these galaxies.}
\label{table:roi}
\end{deluxetable}

@@

\begin{deluxetable}{lcccccc}
\tablecolumns{7}
\tablewidth{0pc}
\tablecaption{\emph{Fermi}-LAT upper limits for the Coma cluster core.}
\tablehead{
\colhead{Spatial model} & \multicolumn{6}{c}{Flux U.L.}
}
\startdata
Point source & 1.73 & 1.81 & 0.487 & 0.526 & 0.627 & 0.635
\enddata
\end{deluxetable}
